---
title: "Quick Analysis"
format: html
editor: visual
---

# Introduction

## About the data

The data for this analysis were downloaded from arctic data center on October 7, 2024: [Arctic Data Center](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)

## Setup

Attach important packages

```{r}
library(readr) # functions for reading in types of file
library(here)
```

## Read in the data

```{r}
#| message: false
#| warning: false
bg_chem <- read_csv(here("data","BGchem2008data.csv"))
# dbl refers to number with a decimel in it
# dttm is a date time class

```

# Analysis

## Calculate the summary statistics

```{r}
#| eval: false ### don't run the code in final product
#| echo: false ### don't show the code in the final report either
### print the column names
# shift command c shortcut for commenting out codes. triple hashtags for titles and notes

colnames(bg_chem)

### get the general structure
str(bg_chem)

### first six lines
head(bg_chem)

### get a summary of each column
summary(bg_chem)

### unique values of a column
unique(bg_chem$Date)
```

Calculate the summary statisitcs for nitrate, nitrite, ammonia, and phosphate

```{r}
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
ammonia <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)
# is there a shortcut for running whole code chunk

```

## Calculate the mean Redfield ratio

Redfield ratio is the ratio of nitrogen (added together) to phosphorous

```{r}
ratio <- (nitrate + nitrite + ammonia)/phos
```

The redfield ratio for this dataset is approximately `r round(ratio)`
Hitting command + return within those ticks tells us what that number is going to look like (i.e. "6")

## Plot Redfield ratio

```{r}
#| echo: false
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```

# Conclusion

